---
title: "FML_2"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2023-10-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Summary

## Questions - Answers

1. How would this customer be classified? This new customer would be classified as 0, does not take the personal loan
2. The best K is 3

## Problem Statement

Universal bank is a young bank growing rapidly in terms of overall customer acquisition.
The majority of these customers are liability customers (depositors) with varying sizes of relationship with the bank. The customer base of asset customers (borrowers) is quite
small, and the bank is interested in expanding this base rapidly in more loan business. In particular, it wants to explore ways of converting its liability customers to personal loan customers.

A campaign that the bank ran last year for liability customers showed a healthy conversion rate of over 9% success. This has encouraged the retail marketing department to devise smarter campaigns with better target marketing. The goal is to use k-NN to predict whether a new customer will accept a loan offer. This will serve as the basis for the design of a new campaign.

The file UniversalBank.csv contains data on 5000 customers. The data include customer
demographic information (age, income, etc.), the customer’s relationship with the bank
(mortgage, securities account, etc.), and the customer response to the last personal loan
campaign (Personal Loan). Among these 5000 customers, only 480 (= 9.6%) accepted the
personal loan that was offered to them in the earlier campaign.

Partition the data into training (60%) and validation (40%) sets

### Data Import and Cleaning

First, load the required libraries
```{r}
library(class)
library(caret)
library(e1071)
```

#Read the data.
```{r}
univers_bank <- read.csv("C:/Users/pardh/Downloads/UniversalBank.csv")

dim(univers_bank)

t(t(names(univers_bank)))
```
```{r}
#Drop ID and ZIP
univers_bank <- univers_bank[, -c(1, 5)]

```

# Take Education as a factor (categorical Predictor) and coverting into dummy variables
```{r}
univers_bank$Education <- as.factor(univers_bank$Education)

groups <- dummyVars(~., data = univers_bank)

univers_bank_m <- as.data.frame(predict(groups, univers_bank))


```

#Split Data into 60% training and 40% validation. There are many ways to do this. We will look at 2 different ways. Before we split, let us transform categorical variables into dummy variables
```{r}
set.seed(123)

train_d <- sample(row.names(univers_bank_m), 0.6*dim(univers_bank_m)[1])

valid_d <- setdiff(row.names(univers_bank_m), train_d)

train_dn <- univers_bank_m[train_d, ]

valid_dn <- univers_bank_m[valid_d, ]

t(t(names(train_dn)))
```


```{r}
#Now, let us normalize the data
train_normald <- train_dn[, -10]

valid_normald <- valid_dn[, -10]

normvalues <- preProcess(train_dn[, -10], method = c("center", "scale"))

train_normald <- predict(normvalues, train_dn[, -10])

valid_normald <- predict(normvalues, valid_dn[, -10])
```

# data frame the given data
```{r}
new_customer <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)

new_custom_normd <- new_customer

new_custom_normd <- predict(normvalues, new_customer)
```


#1. K-NN classification for the above data
```{r}
#Now, let us predict using knn

knn_predcn1 <- class::knn(train = train_normald,
                          test = new_custom_normd,
                          cl = train_dn$Personal.Loan, k=1)

knn_predcn1

accuracy <- data.frame(k = seq(1, 15, 1), overallaccuracy = rep(0, 15))

for (i in 1:15) {
  knn_predcn2 <- class::knn(train = train_normald,
                            test = valid_normald,
                            cl = train_dn$Personal.Loan, k = i)
  
  accuracy[i, 2] <- confusionMatrix(knn_predcn2, as.factor(valid_dn$Personal.Loan), positive = "1")$overall[1]
}

which(accuracy[, 2] == max(accuracy[, 2]))

plot(accuracy$k, accuracy$overallaccuracy, pch = 20, col = "black")
```


#Running the confusion matrix for the valid data set using the best K, (k = 3)
```{r}
knn_predcn3 <- class::knn(train = train_normald,
                          test = valid_normald,
                          cl = train_dn$Personal.Loan, k = 3)
knn_predcn3

confusionMatrix(knn_predcn3, as.factor(valid_dn$Personal.Loan), positive = "1")

```


#4. Consider the following customer: Age = 40, Experience = 10, Income = 84,
Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0,
Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit
Card = 1. Classify the customer using the best k.

```{r}
new_customer1 <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)

new_custom_normd1 <- new_customer1

new_custom_normd1 <- predict(normvalues, new_custom_normd1)

knn_predcn4 <- class::knn(train = train_normald, 
                          test = new_custom_normd1,
                          cl = train_dn$Personal.Loan, k = 3)

knn_predcn4


```

# Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply the k-NN method with the k chosen above. Compare the confusion matrix of the test set with that of the training and validation sets. Comment on the differences and their reason
```{r}
set.seed(2)

train_d1 <- sample(row.names(univers_bank_m), 0.5*dim(univers_bank_m)[1])
train_dn1 <- univers_bank_m[train_d1, ]

valid_d1 <- setdiff(row.names(univers_bank_m), train_d1)
valid_dn1 <- univers_bank_m[valid_d1, ]

valid_d2 <- sample(row.names(valid_dn1), 0.6*dim(valid_dn1)[1])
valid_dn2 <- valid_dn1[valid_d2, ]

test_d1 <- setdiff(row.names(valid_dn1), valid_d2)
test_dn1 <- valid_dn1[test_d1, ]
```

# Normalize the above data sets
```{r}
train_normald1 <- train_dn1[, -10]

valid_normald2 <- valid_dn2[, -10]

test_normal1 <- test_dn1[, -10]

normvalues1 <- preProcess(train_dn1[, -10], method = c("center", "scale"))

train_normald1 <- predict(normvalues1, train_dn1[, -10])

valid_normald2 <- predict(normvalues1, valid_dn2[, -10])

test_normal1 <- predict(normvalues1, test_dn1[, -10])

```

#Knn for training data (50%)
```{r}
knn_predcn5 <- class::knn(train = train_normald1,
                          test = train_normald1,
                          cl = train_dn1$Personal.Loan, k =3)
knn_predcn5

confusion_matrix1 <- confusionMatrix(knn_predcn5, as.factor(train_dn1$Personal.Loan))

confusion_matrix1
```

#Knn for validation data (30%)
```{r}

knn_predcn6 <- class::knn(train = train_normald1,
                          test = valid_normald2,
                          cl = train_dn1$Personal.Loan, k =3)
knn_predcn6

confusion_matrix3 <- confusionMatrix(knn_predcn6,as.factor(valid_dn2$Personal.Loan))
confusion_matrix3
```
#Knn for test data (20%)
```{r}
knn_predcn7 <- class::knn(train_normald1,
                          test = test_normal1,
                          cl = train_dn1$Personal.Loan, k = 3)

knn_predcn7

confusion_matrix4 <- confusionMatrix(knn_predcn7, as.factor(test_dn1$Personal.Loan))

confusion_matrix4
```
Reason: Comparing the confusion matrices of test data with that of training and validation data, the training data is higher compared to the test and validation data respectively. This indicates that the algorithm is running correct.
  